\large Neural and Symbolic Optimization of Cosmological Particle-Mesh Simulation \vspace{0.3em}

### 2023-09-13 docs/papers/sto/so.tex
FIXUP draft updating

\newcommand{\pmwd}{{\usefont{T1}{nova}{m}{sl}pmwd}}
\newcommand{\GADGET}{{{\fontsize{10pt}{12pt}\selectfont GADGET}-4}}

@@

We generate our target data for training and testing with the \GADGET\
simulation code \citep{GADGET-4}.
We run 512 \GADGET\ simulations that cover a wide range of
configurations and cosmological parameters, as listed in
\autoref{tab:param}.
Each simulation has $128^3$ particles and writes 121 snapshots linearly
spaced in scale factor from $1/16 + \Delta a$ to $1 + \Delta a$, with
spacing $1/128$.
$\Delta a$ is an offset, common for all snapshots in each simulation,
that varies as in \autoref{tab:param} and helps to sample different
alignment between snapshot and start/stop times.
In our \pmwd\ model, we generate the snapshot at an arbitrary given time using
the cubic Hermite interpolation, with more details in \autoref{sec:snapobs}.

The curvature $\Omega_k$ is related to the separate universe simulation
\citep{LiEtAl2014, WagnerEtAl2015}.

@@

\vspace{1em}
\subsection{Loss Function}

@@

Following \autoref{tab:param}, in each iteration, we randomly sample the number
of time steps and the mesh shape (or its cell size), and configure \pmwd\ with
the same initial coniditions, cosmological parameters and other common
configurations.
To balance the work of GPUs during training, in each iteration we use
the same number of time steps and mesh shape on all GPUs.

The loss and gradients are collected and averaged across the GPU devices.
The SO parameters are initialized to be the same for all devices, and being
updated with the same averaged gradients.

@@

Hyperparameters:
\YZ{
\begin{itemize}
  \item Model: number of hidden layers and nodes of f and g nets, activation function
  \item Optimization: learning rate scheduler, momentum, weight decay and dropout rate etc
  \item Loss: fields to include and loss function forms
\end{itemize}
}

@@

\YZ{
We implement the neural networks with \texttt{Flax} \citep{flax2020github}, and
use the optimizers in \texttt{Optax}, part of the DeepMind JAX Ecosystem
\citep{deepmind2020jax}.
}

@@

\vspace{1em}
\textit{\large Data Availability:}
\YZ{Do we give \GADGET\ Sobol 512 dataset a name and make it public? This could
be useful for future training projects.}

>>>>

\newcommand{\pmwd}{{\usefont{T1}{nova}{m}{sl}pmwd}}
\newcommand{\GADGET}{{{\fontsize{10pt}{12pt}\selectfont GADGET}-4}}
\newcommand{\GSDATA}{{GS512}}

@@

We generate our target data for training and testing with the \GADGET\
simulation code \citep{GADGET-4}.
We run 512 \GADGET\ simulations that cover a wide range of
configurations and cosmological parameters, as listed in
\autoref{tab:param}.
Each simulation has $128^3$ particles and writes 121 snapshots linearly
spaced in scale factor from $1/16 + \Delta a$ to $1 + \Delta a$, with
spacing $1/128$.
$\Delta a$ is an offset, common for all snapshots in each simulation,
that varies as in \autoref{tab:param} and helps to sample different
alignment between snapshot and start/stop times.
In our \pmwd\ model, we generate the snapshot at an arbitrary given time using
the cubic Hermite interpolation, with more details in \autoref{sec:snapobs}.
The curvature $\Omega_k$ is related to the separate universe simulation
\citep{LiEtAl2014, WagnerEtAl2015}.
\YZ{We refer to this training dataset of 512 \GADGET\ simulations sampled with
Sobol sequence as \GSDATA.}

@@

\vspace{1em}
\subsection{Loss Function}
\label{sec:loss}

@@

Following \autoref{tab:param}, in each iteration, we randomly sample the number
of time steps and the mesh shape, and configure \pmwd\ with the same initial
coniditions, cosmological parameters and other common configurations.
To balance the work of GPUs during training, in each iteration we use
the same number of time steps and mesh shape on all GPUs.

As discussed in \autoref{sec:loss}, the loss in a single simulation sums over
all snapshots.
Our batch size is [...], i.e. the number of GPU devices since one simulation is
run on a given GPU and compared to the corresponding training data sample for
loss.
For each training step, the loss and gradients are collected and averaged across
all GPU devices.
The SO parameters are initialized to be the same for all devices, and being
updated with the same averaged gradients.

@@

Hyperparameters:
\begin{itemize}
  \item Model: number of hidden layers and nodes of f and g nets, activation function
  \item Optimization: learning rate scheduler, momentum, weight decay and dropout rate etc
  \item Loss: fields to include and loss function forms
\end{itemize}

@@

We implement the neural networks with \texttt{Flax} \citep{flax2020github}, and
use the optimizers in \texttt{Optax}, part of the DeepMind JAX Ecosystem
\citep{deepmind2020jax}.

@@

\vspace{1em}
\textit{\large Data Availability:}
\YZ{Do we make the \GSDATA\ data public?}
